The Countercyclical Component of Capital Regulation

### 2020-04-14 Analysis/Draft/Draft_bookdown.Rmd
Start portfolio reallocation section

source("../../Auxiliar/Function.R") ## Capital requirements functions
source("../../Auxiliar/longtable.R") ## longtable for stargazer
source("../../Auxiliar/adj_stargazer.R") ## adjust the first column width for stargazer

@@

The results from Tables \ref{tab:main} and \ref{tab:alt} raises the question of why it appears that only *Capital savings* from corporate portfolios are impacted during the business cycle. In this section I argue that portfolio reallocation during the business cycle explains part of *Capital savings* cyclicality. Ideally, for this purpose, I would track individual assets during the cycle and compare their PDs evolution to the PDs of new loans.  Unfortunately, I can only observe assets aggregated by PD bands which does not allow me to track individual assets. As an alternative approach, I analyze the relationship of PD distribution moments with GDP to understand asset allocation during the business cycle.

According to the theoretical framework presented in section 4.1, an increase in average PD reduces the amount of *Capital savings* if other moments are kept constant. This relationship is illustrated in panel A of Figure \ref{fig:pdmapcases} and empirically tested in Table \ref{tab:moments_contribution}. In all columns of this table average PD is negatively associated with *Capital savings*. Table \ref{tab:moments_contribution} also shows that, in line with the theoretical framework, an increase in PD variance is positively associated with *Capital savings* for all portfolios. In sum, the estimated coefficients of the average PD and the PD variance provide strong empirical support to the theoretical framework presented in section 4.1. At the same time, however, I cannot refute the hypothesis that portfolio reallocation does not impact *Capital savings* because the estimated effect of these two moments on *Capital savings* are in line with a scenario without any portfolio reallocation.

Extending the analysis to the skewness and kurtosis of the PD distribution proves to be very helpful. Consider a

Panel A in table \ref{tab:moments_cyclicality} reports how average PD moves during the business cycle. As expected, in most of the regressions, changes in average PD are negatively associated with GDP growth. This association is the known driver of the procyclical character of the current model-based capital regulation and its interpretation is fairly mechanical. During recessions average credit quality deteriorates, or in other words, average PD increases. The only exception to the rule is the sovereign portfolio. Column 5 of panel A shows that the coefficient of GDP growth on changes on average PD of sovereign portfolios is positive, although without strong evidence that it is not different from zero. One possible reason for this result is that, during economic downturns, banks may search for safe assets mostly in the form of safe haven government bonds. According to column 5 of panel A, this reallocation of portfolio appears to be sufficient to counterbalance the deteriorations of the prior portfolio. It also helps explain the insignificant results in column 5 of Tables \ref{tab:main} and \ref{tab:alt}.

As explained, if no other moment is affected during the business cycle we should observe procyclical *Capital savings*. However, it was shown in

Common  
I further analyse the relationship of the PD distribution moments with GDP and find that during economic expansions: average, variance and kurtosis decrease while the skewness increases. This evidence is in line with the idea that the upper-tail (riskier borrowers) of the PD distribution is disproportionately affected by the business cycle.

@@

The follow-up to the 2008 financial crisis offer an plausible explanation for the dynamics of *Capital savings* in time, which is illustrated in Figure \label{fig:timevariables}. Increased supervisory pressure to increase capital ratios (top plot), especially after the Basell III, led banks to deleverage and reduce risk-taking. [^1] The shift for safer assets not only reduced the average PD on banks' portfolios (middle plot) but also compressed their PD distribution (bottom plot). This distribution compression leads to a reduction in *Capital savings*. This explanation is in line with the hypothesis that *Capital savings* are countercyclical. Efforts to increase capital ratio by risk reduction are increasingly more difficult.

[^1]: Although supervisors favored raising additional capital as means to achieve the required capital ratios [@EBA2011recommendation], banks respond to capital requirements mostly by reducing RWA [@gropp2018banks].

>>>>

source("../../Auxiliar/Function.R") ## Capital requirements functions
source("../../Auxiliar/longtable.R") ## longtable for stargazer
#source("../../Auxiliar/adj_stargazer.R") ## adjust the first column width for stargazer

@@

In this section I argue that portfolio reallocation during the business cycle explains part of the *Capital savings* cyclicality. As method, I use the observed movements on moments of the PD distribution during the business cycle that are not in line with the expected movements from a static portfolio as evidence in favor of a portfolio reallocation effect on *Capital savings*.[^1]

[^1]: Ideally, for this purpose, I would track individual assets during the cycle and compare their PDs evolution to the PDs of new loans. Unfortunately, in my dataset, I only observe assets aggregated by PD bands without knowledge if amount changes in a band are due to new loans or changes in credit quality of assets already held.

According to the theoretical framework presented in section 4.1, an increase in average PD reduces the amount of *Capital savings* if other moments are kept constant. This relationship is illustrated in panel A of Figure \ref{fig:pdmapcases} and empirically tested in Table \ref{tab:moments_contribution}. In all columns of this table average PD is negatively associated with *Capital savings*.

Panel A in table \ref{tab:moments_cyclicality} reports how average PD moves during the business cycle. As expected, in most of the regressions, changes in average PD are negatively associated with GDP growth. This association is the known driver of the procyclical character of the current model-based capital regulation and its interpretation is fairly mechanical. During recessions average credit quality deteriorates, or in other words, average PD increases. The only exception to the rule is the sovereign portfolio. Column 5 of panel A shows that the coefficient of GDP growth rate on changes in the average PD of sovereign portfolios is positive, although without strong evidence that it is not different from zero. One possible reason for this result is that, during economic downturns, banks may search for safe assets mostly in the form of safe haven government bonds. According to column 5 of panel A, this reallocation of portfolio appears to be sufficient to compensate any deterioration of the prior held portfolio. Together with column 5 of Table \ref{tab:moments_contribution}, it also helps explain the positive (and insignificant) results in column 5 of Tables \ref{tab:main} and \ref{tab:alt}.

Table \ref{tab:moments_contribution} also shows that, in line with the theoretical framework, an increase in PD variance is positively associated with *Capital savings* for all portfolios. And in line with the evidence from Tables \ref{tab:main} and \ref{tab:alt}, Panel B in table \ref{tab:moments_cyclicality} shows that the effect of the business cycle on the PD variance is concentrated on corporate portfolios. In sum, the estimated coefficients of the average PD and the PD variance in Table \ref{tab:moments_contribution} and panels A and B in Table \ref{tab:moments_cyclicality} provide strong empirical support to the theoretical framework presented in section 4.1.

At the same time, however, using only the evidence provided by the average and variance of the PD distribution, the hypothesis that portfolio reallocation does not impact *Capital savings* cannot be refuted exactly because the estimated effect of these two moments on *Capital savings* are strongly in line with a model without any portfolio reallocation. Consider now a negative shock to a continuous PD distribution, such that we can measure its skewness and kurtosis. It is trivial that the equivalent scenario of panel A of figure \ref{fig:pdmapcases} wouldn't affect either of these moments. On the other hand, in the second scenario the PD distribution should become more skewed to the right. Even if we think that some assets will become safer during bad times, it is reasonable to expect a larger effect on the right tail than on the left tail of the PD distribution. Consequently, considering a static portfolio, we expect the skewness of the PD distribution to be negatively associated with the GDP growth rate.

Panel C in table \ref{tab:moments_cyclicality} provides evidence that the observed relationship of PD distribution skewness and the business cycle can only be explained with a within portfolion reallocation.

@@

The follow-up to the 2008 financial crisis offer an plausible explanation for the dynamics of *Capital savings* in time, which is illustrated in Figure \label{fig:timevariables}. Increased supervisory pressure to increase capital ratios (top plot), especially after the Basell III, led banks to deleverage and reduce risk-taking. [^2] The shift for safer assets not only reduced the average PD on banks' portfolios (middle plot) but also compressed their PD distribution (bottom plot). This distribution compression leads to a reduction in *Capital savings*. This explanation is in line with the hypothesis that *Capital savings* are countercyclical. Efforts to increase capital ratio by risk reduction are increasingly more difficult.

[^2]: Although supervisors favored raising additional capital as means to achieve the required capital ratios [@EBA2011recommendation], banks respond to capital requirements mostly by reducing RWA [@gropp2018banks].
